Potential for remotely sensing riverine gas exchange and carbon efflux from SWOT observations

### 2021-05-24 manuscript_v3.Rmd
big ms edits

##Highlights (3-5 points, 85 characters each w/ spaces)
- The SWOT satellite should be able to indirectly remotely sense riverine gas exchange velocity
- The BIKER algorithm predicts $CO_2$ fluxes when coupled with in situ gas concentration data
- Predicted BIKER $CO_2$ fluxes compare well with in-situ methods
- Expected SWOT measurement errors exert a marginal influence on BIKER's estimates
- Errors are dominated by process-level uncertainties, not parameter inversion

@@

##Abstract
Go BIKER!!
max 400 words
x
x
x
x
x
x
x
x
x

##1 Introduction
Inland waters transport and transform various constiutents as they move from the landscape to the ocean [@colePlumbingGlobalCarbon2007a]. **Too big a jump. What is the journal? ‘constituents’ will need to be defined, as will ‘processes’**As a result some of these processes, rivers and lakes are usually supersaturated with dissolved greenhouse gases (GHGs) and thus present a significant flux of these gases from water to air [@raymondGlobalCarbonDioxide2013a]. This flux $[M/L^2T]$ is expressed as equation 1 given the gas concentration gradient between the water $C_{water} [M/L^3]$ and the air $C_{air} [M/L^3]$ and the gas exchange velocity $k [L/T]$. Note that *k* is generally studied as normalized by a Schmidt number of 600 ($k_{600}$) to be gas and temperature agnostic [see @hallGasExchangeStreams2020 for a thorough review of gas exchange in rivers].

@@

To directly measure $k_{600}$ in a given stream, scientists use a range of methods mostly involving floating domes or eddy covariance towers to measure at a point in the river, or using injections of tracer gases to estimate reach-scale gas exchange [again, see @hallGasExchangeStreams2020]. Obiously, these types of of measuremnets are preferred but are infeasible when working at the network-scale across potentially tens of thousands of rivers, and so upscaling functions are used to predict $k_{600}$ from readily available river geomorphology parameters [e.g. @oconnorMechanismReaerationNatural1958; @palumboAssessingPerformanceReaeration2014; @raymondScalingGasTransfer2012a]. Many upscaling functions are based on the basic theory that *k* is empirically correlated with the turbulent energy dissipation *e* at the air-water interface [@zappaEnvironmentalTurbulentMixing2007]. This is convenient for upscaling because in rivers and streams *e* can be equated to the turbulent energy dissipation rate *eD* $[L^2/T^3]$ [@ulsethDistinctAirWater2019a; @raymondScalingGasTransfer2012a], which is estimated using easily calculated hydraulic parameters: $eD \approx gSV$ where *g* is gravitational acceleration $[L/T^2]$, *S* is channel slope $[L/L]$, and *V* is average flow velocity $[L/T]$ [@tsivoglouTracerMeasurementReaeration1976]. Because channel slope is readily available in any hydrographic data product, most efforts to upscale $k_{600}$ to drainage networks are thereby limited by the quality of the final parameter *V*. Since *V* must be measured in the field or itself estimated by other models/empirical methods, uncertainty in *V* must dominates errors in $k_{600}$ estimates. This is exacerbated in ungauged basins that cover large sections of the earth system, especially in the carbon-rich Arctic inland waters, as no in situ hydraulic information is available and fieldwork is impractical [@gleasonRemoteSensingRiver2020].

To circumvent this problem in ungauged basins and at network-scales, researchers have used ‘global-scope’ hydraulic geometry (HG) models: simple power law relations between in this case *V* and discharge that are trained on extensive hydraulics datasets [@leopoldHydraulicGeometryStream1953a; @borgesVariabilityGasTransfer2004; @horgbyUnexpectedLargeEvasion2019; @lauerwaldSpatialPatternsCO22015; @raymondGlobalCarbonDioxide2013a]. However, it has long been established that HG parameters are highly variable across landscapes and from river to river [@gleasonHydraulicGeometryNatural2015; @parkWorldwideVariationsHydraulic1977; @rhodesBfmDiagramGraphical1977]. Because of this, a large body of geomorphology work has attempted to parse out process-based explanations for HG parameters [e.g. @dingmanAnalyticalDerivationAtastation2007b; @parkerPhysicalBasisQuasiuniversal2007b; @singhTHEORIESHYDRAULICGEOMETRY2003a; @fergusonHydraulicsHydraulicGeometry1986a] and an open research question is how to best prescribe a set of HG parameters to a given river (particularly at the global scale). **Xxx give me a sentence or two on the problems with this is approach here. You bury the problem with the sentence “however, it has long” but then cover it up. The transition doesn’t work without closing this paragraph with a supported (i.e. with citations) statement explicitly defining why we can’t do the global scope HG approach. xxx **

A potential alternative to this ‘global-scope’ HG approach is to directly estimate a river’s hydraulic properties (e.g. V) from remote sensing (RS) data. Remote sensing of river hydraulics is a burgeoning subfield within remote sensing of hydrology, typically in service of remote sensing of river discharge (RSQ- see @gleasonRemoteSensingRiver2020 for a thorough review). There are many ways to perform RSQ, from calibrating RS data to local channel hydraulics [e.g. @brakenridgeOrbitalMicrowaveMeasurement2007; @pavelskyUsingWidthbasedRating2014; @pavelskyRemoteSensingSuspended2009; @tarpanelliEstimationRiverDischarge2013] to calibrating hydraulic/hydrologic models with both in situ and RS data [e.g. @bjerklieEstimatingDischargeRivers2005; @chandanpurkarSatelliteReanalysisBasedMass2017; @linGlobalReconstructionNaturalized2019a; @nealDataAssimilationApproach2009] to methods that use no in situ information in their hydraulic/hydrologic models (termed 'ungauged approaches'). Many, but not all, of these ungauged approaches were developed in the context of the upcoming NASA/CNES/UKSA/CSA Surface Water and Ocean Topography (SWOT) satellite mission. **CRAIG: might need to add unaguged citations here or rewrite perspective**

SWOT is expected to launch in 2022 and provide the world’s first global measurements of water surface extent and elevation at novel temporal resolutions. SWOT is a wide swath Ka-band radar interferometer and will sample rivers every 1 to 7 days per 21 day repeat cycle that will measure rivers wider than 100m with a goal of expanding this to rivers at least 50m wide [@biancamariaSWOTMissionIts2016]. A decade of SWOT work has explored the multi-parameter problem of estimating the river hydraulic parameters of roughness and bathymetry from remote sensing to produce the SWOT discharge product [e.g. @andreadisConstrainingAssimilationSWOT2020d; @brinkerhoffConstrainingRemoteRiver2020; @durandEstimatingReachaveragedDischarge2014a; @garamboisInferenceEffectiveRiver2015; @gleasonRetrievalRiverDischarge2014a; @garamboisVariationalEstimationEffective2020; @hagemannBAMBayesianAMHGManning2017a; @larnierRiverDischargeBathymetry2020a; @oubanasDischargeEstimationUngauged2018]. A specific subset of these ungauged methods are termed ‘McFLIs’ or ‘Mass Conserved Flow Law Inversion’ algorithms [e.g. @andreadisConstrainingAssimilationSWOT2020d; @brinkerhoffConstrainingRemoteRiver2020; @hagemannBAMBayesianAMHGManning2017a; @durandEstimatingReachaveragedDischarge2014a]. These use basic geomorphic theories and the concept of ‘prior river knowledge’ to estimate discharge from RS data where not all hydraulic terms are RS-able [@gleasonRemoteSensingRiver2020; @gleasonTrackingRiverFlows2017]. McFLIs are globally flexible, easily implementable in any river that SWOT can sample, and improve our hydrological understandings of ungauged rivers when little to no information was previously available [@brinkerhoffConstrainingRemoteRiver2020; @durandIntercomparisonRemoteSensing2016b]. Recently, McFLIs have also shown promise in providing additional and beneficial information to traditional hydrologic modeling via data assimilation [@ishitsukaCombiningOpticalRemote2020].

While all McFLIs to date have been developed in the context of RSQ, there is no reason their logic cannot be used to estimate $k_{600}$, particularly as McFLIs often employ Bayesian inference for equifinal inverse problems. Equifinality refers to an under-constrained mathematical system that have essentially infinite parameter combinations that can produce the same result: there are in essence more unknowns than equations [@garamboisInferenceEffectiveRiver2015], and this problem is common to both RSQ and $K_{600}$ estimation. Bayesian techniques have been previously used to estimate $k_{600}$ concurrently with stream metabolism from dissolved oxygen ($O_2$) datasets [@applingOvercomingEquifinalityLeveraging2018; @graceFastProcessingDiel2015; @holtgrieveSimultaneousQuantificationAquatic2010]. While these studies are field based and are not applicable to $k_{600}$ estimation in ungauged rivers, they suggest that Bayesian techniques could be useful for remotely sensing $k_{600}$ when parameter equifinality is a problem, which matches McFLI logic precisely [@andreadisConstrainingAssimilationSWOT2020d].

In this context, we hypothesize that a combination of SWOT data, a Bayesian McFLI paradigm, and gas exchange theory can successfully estimate $K_{600}$. This would in turn improve our understanding of riverine GHG fluxes in ungauged basins. Therefore, we developed a method to estimate $k_{600}$ solely from SWOT measurements, using hydrologic model data to produce SWOT-like data as SWOT has not yet launched [standard practice in the SWOT community- e,g, @durandIntercomparisonRemoteSensing2016b]. We hypothesize that remote sensing $k_{600}$ is possible by coupling Bayesian remote sensing techniques (in a manner similar to McFLI) and SWOT data with hydraulic geometry and gas exchange theory. More specifically, this manuscript aims to answer two questions:

- Is ungauged RS of the gas exchange velocity ($K_{600}$) possible using soon-to-be-available SWOT river measurements?
- If possible, how will this RS method change estimates of riverine carbon efflux at network scales?

To answer these questions, we developed a new McFLI algorithm that ingests SWOT data and produces $k_{600}$ estimates and its explicit Bayesian uncertainity that requires no in situ inputs of any kind (although in situ data can be ingested and will improve results). We name the RS of $k_{600}$ algorithm BIKER, or the ‘Bayesian Inversion of the $k_{600}$ Rate’. We validate BIKER on simulated SWOT data for 22 rivers. We also quantify BIKER's sensitivity to the expected SWOT measurement errors: while SWOT data represent a sea change in inland water monitoring, SWOT is expected to have an approximately 10cm error in water surface elevation [@biancamariaSWOTMissionIts2016]. Finally, we use previously published dissolved $CO_2$ data to compare the bulk carbon efflux (via $CO_2$ evasion) from the 22 rivers as calculated using BIKER and previously published in situ techniques to show how BIKER’s $k_{600}$ estimates translate into carbon efflux.

##2 Methods
To build BIKER, we join a process-based model for $k_{o2}$ with a McFLI framework for inverting SWOT measurements (section 2.1). Following the description of that process, we describe the validation (section 2.3) and the workflow for comparing estimated bulk carbon effluxes from a suite of hydraulic geometry models (section 2.4). A flowchart detailing the entire study is given as Figure 1.

![Figure 1. Flowchart of methodology used in this study. We begin by deriving a process-based model for ko2 that uses only river width and slope to assign its parameters (section 2.1). Then, we implement this model within a McFLI ('BIKER') to estimate ko2 solely from river width and water surface slope (section 2.2). Finally, we couple BIKER with in situ CO2 data to compare BIKER-estimated gas fluxes with established in situ methods (section 2.4). See section 2.3 for the validation setup.](cache/flowchart.jpg)

###2.1 Developing BIKER
To use the McFLI framework, a physical model for the parameter of interest is necessary. In this case we need a physical model for *k*. First, we detail which model for *k* we chose to implement, and then we explain how the implementation works.

There is a long tradition of water quality engineers and biogeochemists developing predictive models for different forms of *k*, going back to at least the 1950s [@wangPhysicallyBasedScaling2021]. The goals of these models vary, from predicting oxygen reaeration coefficients to predicting $CO_2$ gas transfer velocities. Most of these models operate under the basic assumption that river channel hydraulics govern gas exchange and so simple statisical relationships can be generated between forms of *k* and hydraulic properties. However, many of them are not rooted in process-based mass transport theory and rather are simply empirical relations. This means that many of these models exist at the mercy of the limited training data available to fit said models and are highly sensitive to which data are used.

Recently, many of these models were rexplored in rivers and streams by @wangPhysicallyBasedScaling2021. They significantly expanded training datasets of field-measured $k_{o_2}$ (*k* for dissolved oxygen DO at 20 degrees Celsius) by using inverse modeling of DO timeseries. This apprroach was briefly mentioned in the introduction of this manuscript. These methods calibrate $k_{o_2}$, along with respiration and primary production, to high-fidelity, in situ datasets of DO with the goal of minimizing the difference between the observed DO and the mass-balance DO [e.g. @applingOvercomingEquifinalityLeveraging2018; @chapraDeltaMethodEstimating1991]. @wangPhysicallyBasedScaling2021 used this technique at 35 rivers across the United States and showed that $k_{o_2}$ models deduced from the classic, process-based 'thin-film' and 'surface renewal' theories yield approximately identical parameters when fit to field measurements or this new dataset of simulated $k_{o_2}$. This suggests that there are uniform scaling realtionships between certain hydraulic properties and $k_{o_2}$ that exist regardless of the training data used. See [@hallGasExchangeStreams2020 and @wangPhysicallyBasedScaling2021] for more in-depth explanations of the thin-film and surface renewal theories with respect to rivers. These are the standard models for understanding gas exchange in many aquatic environments from lakes to esutraries to rivers. The better fit, and more parsimonious, of these two models is reprinted as equations 2 and 3 where $U_\star$ is the shear velocity of the river channel [L/T] ($U_\star = \sqrt{gR_hS_h}$). Equation 2 was fit to field measurements of $k_{o_2}$ while equation 3 was fit to the simulated $k_{o_2}$ for 35 rivers.
$$k_{o2} = 47.67U_\star \mathbf{(2)}$$
$$k_{o2} = 49.71U_\star \mathbf{(3)}$$

This model was implemented in BIKER (using 48 as the approximate parameter across equations 2 and 3) for a few reasons. First, the analysis performed by [@wangPhysicallyBasedScaling2021] provides compelling evidence that a statistical parameter of approximately 48 is uniform across rivers and streams, regardless of the training data used. This is encouraging for use within BIKER as BIKER is specifically designed to be as river-agnostic as possible. Second, it yields a simple linear relationship across all rivers, regardless of their size or steepeness. This is not necessairly true of *k* models based on the turbulent dissipation rate *eD* ($eD=gS_hV$), which is another common approach [@hallGasExchangeStreams2020; e.g. @raymondScalingGasTransfer2012a; @oconnorMechanismReaerationNatural1958; @tsivoglouTracerMeasurementReaeration1976], where recent work has shown that *k* does not scale with $eD$ via the same statistical parameters across all rivers [@ulsethDistinctAirWater2019a]. We further explored the relationship presented in that paper and found that the $k~eD$ relationship fundamentally breaks down in rivers with low *eD*, which corresponds to nearly all rivers that SWOT will observe (**supp figure**). Therefore, an $eD$ based predictive model is of little use in SWOT rivers. Third, we sought to explicitly predict the gas transfer velocity for DO, as its use extends beyond constraining GHG fluxes from rivers. For example **stuff here**. To convert from $k_o2$ to *k* for any other GHG (as will be done later in this manuscript to calculate $CO_2$ evasion fluxes), we use the Schmidt number to convert between water temperatures and gases of interest following equation 4. 530 is the Schmidt number for oxygen at 20 degrees Celsius and *Sc* is the Schmidt number at one's desired temperature for some dissolved gas.
$$k_{GHG, temp}=k_{02}*(Sc/530)^{-1/2} \mathbf{(4)}$$

We stress that most other predictive models for *k* could be implemented within BIKER. However, our goal of using BIKER on any river that is observable by SWOT necessitates that we pick an equation that is applicable across as many rivers as possible. Future work should look at using a river-specific *k* equation to 'calibrate' BIKER to a specific river of interest.

With a predictive model for $k_{o2}$ chosen, we now implement it within a McFLI framework. The approach used here is strongly informed by the @hagemannBAMBayesianAMHGManning2017a McFLI algorithm for ungauged RSQ, further explored in more recent work [@brinkerhoffConstrainingRemoteRiver2020]. These papers conceptualize discharge as a Bayesian remote sensing problem, which we largely follow here to conceptualize $k_{o2}$ as a Bayesian remote sensing problem that can be solved using SWOT data.

First, the model needs to be rewritten as a function of SWOT-observables river width *W* and height (and thus water surface slope $S_h$). SWOT will measure these at approximately 200m intervals in pre-defined mass-conserved river reaches (**What is the PDD?**; SWORD). To do this, we follow @hagemannBAMBayesianAMHGManning2017a to formulate depth via river width and area. Area is further split into the SWOT-observable portion *dA* and the unobservable portion $A_0$ following @durandEstimatingReachaveragedDischarge2014a. $A_0$ is the median channel area and *dA* is the change in channel area. We estimate *dA* by assuming a rectangular river channel so that $dA \approx \delta H*W$ and is also solely a function of SWOT-observables *W* and $S_h$.

**COLIN ABOUT DA: sensitivity analysis? "We investigated the sensitivity of BIKER to this rectangular assumption and found it negligible (xxx section x.x if appropriate xxx)".**

>>>>

##Highlights (3-5 points, 85 characters each w/ spaces)
- The BIKER algorithm indirectly remotely senses riverine gas exchange velocity
- Predicted gas exchange velocities compare well with in-situ methods
- BIKER predicts $CO_2$ fluxes when coupled with in situ gas concentration data
- Expected SWOT measurement errors exert a marginal influence on BIKER's estimates
- Errors are dominated by process-level uncertainties, not parameter inversion

@@

##Abstract
Go BIKER!!
*Note: this is for Remote Sensing of Environment (I think). So 400 words here.*

##1 Introduction
The Earth's carbon cycle transports and manipulates carbon between the atmosphere, geosphere, hydrosphere, and biosphere via a complex set of interconnected biogeochemical processes. The primary connection between the terrestrial and oceanic biospheres is the global river network, which exports much carbon from the continents to the oceans [@colePlumbingGlobalCarbon2007a]. However en route to the oceans, this river network additionally transports and transforms sediments, nutrients, organic matter, dissolved gases, and other constituents. Organic matter and inorganic nutrients are also produced and/or consumed within rivers via ecosystem metabolism [@odumPrimaryProductionFlowing1956]. The net effect of all of this is the global river network is usually supersaturated with dissolved gases and evades substantial amounts of carbon from its water surface to the atmosphere [@colePlumbingGlobalCarbon2007a].

Current estimates of total carbon dioxide ($CO_2$) evasion to the atmosphere vary from 650-1800 Tg C/yr [@lauerwaldSpatialPatternsCO22015; @raymondGlobalCarbonDioxide2013a], with 167 Tg-C/yr coming from mountain streams alone [@horgbyUnexpectedLargeEvasion2019]. Further, in-river metabolism is increasingly understood to contribute to this flux [@hotchkissSourcesProcessesControlling2015]. Equation 1 represents this flux for any sparingly soluble gas $[M/L^2T]$ given the gas concentration gradient between the water $C_{water} [M/L^3]$ and the air $C_{air} [M/L^3]$ and the gas exchange velocity $k [L/T]$. Given the structure of equation 1, calculations of this flux are highly sensitive to our measurements/estimates of *k*. Note that *k* is different for different gases and water temperatures [see @hallGasExchangeStreams2020 for a thorough review of gas exchange in rivers].

@@

*k* is also vital to understanding freshwater ecology through bulk river metabolism [gross primary production and ecological respiration- @bernhardtMetabolicRegimesFlowing2018]. The general mass-balance equation for river metabolism (*M* $[M/L^2T]$) is provided as equation 2, where *DO* is the dissolved oxygen concentration $[M/L^3]$, *D* is river depth [L], and $k_{o_2}$ is *k* for oxygen at some temperature. When viewing equations 1 and 2 in aggregate, it is clear that *k* plays a critical role in two of the fundamental biogeochemical processes that occur in river networks. Therefore, both closing the global carbon cycle and further understanding freshwater ecology rely on accurate and high-fidelity assessments of *k*.
$$\frac{dDO}{dt}=\frac{M}{D}+\frac{k_{o_2}}{D}(DO_{air}-DO_{water}) \mathbf{(2)}$$
Researchers use a range of methods to measure *k*, from floating domes to eddy covariance towers to measure at a specific point in the river. They also inject tracer gases to estimate reach-scale gas exchange [@hallGasExchangeStreams2020]. These approaches are ideal but are infeasible when working at the river network scale across potentially tens of thousands of rivers. So, predictive models are used to estimate *k* from readily available river geomorphology [e.g. @oconnorMechanismReaerationNatural1958; @palumboAssessingPerformanceReaeration2014; @raymondScalingGasTransfer2012a]. These models often, though not alwyas, predict $k_{600}$, or *k* that has been normalized to a Schmidt number of 600 ($CO_2$ at 20 degrees Celsius or $O_2$ at 17.5 degrees).

There is a long tradition of water quality engineers and biogeochemists developing predictive models for different forms of *k*, going back to at least the 1950s [@hallGasExchangeStreams2020] and extending the original work of @streeterMeasuresNaturalOxidation1935 and @phelpsStudiesSelfPurificationStreams1914. The goals of these models vary, from predicting oxygen reaeration coefficients to predicting $CO_2$ gas transfer velocities. Most of these models operate under the basic assumption that river channel hydraulics govern gas exchange and so simple statisical relationships can be generated between *k* and hydraulic properties. However, many of them are not based on fundamental mass transport theory and rather are simple empirical relations. This means that those models' parameters are sensitive to the training data used [@wangPhysicallyBasedScaling2021].

Multiple predictive models are based on the observation that *k* is generally correlated with both the turbulent energy dissipation rate *eD* $[L^2/T^3]$ [@ulsethDistinctAirWater2019a; @raymondScalingGasTransfer2012a] and the water surface surface shear velocity $U_\star [L/T]$ [@wangPhysicallyBasedScaling2021; @katulStructureFunctionModel2018]. This is convenient because both *eD* and $U_\star$ are calculated using easily estimated hydraulic parameters: $eD \approx gS_hV$ [@tsivoglouTracerMeasurementReaeration1976] and $U_\star=\sqrt{gS_hR_h}$, where *g* is gravitational acceleration $[L/T^2]$, $S_h$ is water surface slope $[L/L]$, *V* is average flow velocity $[L/T]$ and $R_h$ is the hydraulic radius [L] (which often approximates *D*). Because slope is readily available in any hydrographic data product, most efforts to predict *k* across thousands of rivers are thereby limited by the quality of the final parameters *V* or $R_h$. Since both must be measured in the field or themselves estimated by other empirical methods, uncertainty in *V* or $R_h$ must dominate any errors in *k* estimates. This is exacerbated in ungauged basins that cover large areas, especially in the carbon-rich Arctic inland waters, as no in situ hydraulic information is available and fieldwork is impractical [@gleasonRemoteSensingRiver2020].

To circumvent this problem in ungauged basins and across thousands of rivers, researchers have used 'global-scope' hydraulic geometry (HG) models that extend HG beyond its traditional use. Traditional HG models are simple power law relations between streamflow and channel width, depth, or velocity at a specific cross-section in a river [@leopoldHydraulicGeometryStream1953a] and a large body of geomorphology work has attempted to parse out process-based explanations for observed HG parameters [e.g. @dingmanAnalyticalDerivationAtastation2007b; @parkerPhysicalBasisQuasiuniversal2007b; @singhTHEORIESHYDRAULICGEOMETRY2003a; @fergusonHydraulicsHydraulicGeometry1986a]. This has been extended to 'downstream HG' which relates a characteristic streamflow (i.e. mean annual or bankful discharge) to characterisitc hydraulics along a specific river's course [@leopoldHydraulicGeometryStream1953a].

For the purposes of predicting *k* from *V* or $R_h$, biogeochemists have further extended HG models by training them on large hydraulics datasets for thousands of rivers and under all streamflow conditions [e.g @borgesGloballySignificantGreenhousegas2015a; @horgbyUnexpectedLargeEvasion2019; @lauerwaldSpatialPatternsCO22015; @raymondGlobalCarbonDioxide2013a]. However, it has long been established that HG parameters are either cross-section specific and highly variable from river to river (traditional HG) or they are discharge specific [downstream HG- @gleasonHydraulicGeometryNatural2015; @parkWorldwideVariationsHydraulic1977; @rhodesBfmDiagramGraphical1977]. This means that the 'global-scope' HG models that are usually used are highly reliant on their training data and can potentially yield pretty different hydraulics for the same streamflow. It is currently not well understood how sensitive global estimates of riverine gas evasion are to the specific 'global-scope' HG model that is employed by the worker.

A potential alternative to this 'global-scope' HG approach is to directly estimate a river's hydraulic properties from remote sensing (RS) data. Remote sensing of river hydraulics is a burgeoning subfield within remote sensing of hydrology, often in service of remote sensing of river discharge (RSQ- see @gleasonRemoteSensingRiver2020 for a recent review). There are many ways to perform RSQ, from calibrating RS data to local channel hydraulics [e.g. @brakenridgeOrbitalMicrowaveMeasurement2007; @pavelskyUsingWidthbasedRating2014; @pavelskyRemoteSensingSuspended2009; @tarpanelliEstimationRiverDischarge2013] to calibrating hydraulic/hydrologic models with both in situ and RS data [e.g. @bjerklieEstimatingDischargeRivers2005; @chandanpurkarSatelliteReanalysisBasedMass2017; @linGlobalReconstructionNaturalized2019a; @nealDataAssimilationApproach2009] to methods that use no in situ information in their hydraulic/hydrologic models (termed 'ungauged approaches' and detailed below). Many, but not all, of these ungauged approaches were developed in the context of the upcoming NASA/CNES/UKSA/CSA Surface Water and Ocean Topography (SWOT) satellite mission.

SWOT is expected to launch in 2022 and provide the world's first global measurements of water surface extent and elevation at novel temporal resolutions. SWOT is a wide swath Ka-band radar interferometer and will sample rivers every 1 to 7 days per 21 day repeat cycle. This yields an average sampling resolution of 11 days. It will measure rivers wider than 100m with a goal of expanding this to rivers at least 50m wide [@biancamariaSWOTMissionIts2016]. A decade of SWOT work has explored the multi-parameter problem of estimating the river hydraulic parameters of roughness and bathymetry from remote sensing to produce the SWOT discharge product [e.g. @andreadisConstrainingAssimilationSWOT2020d; @brinkerhoffConstrainingRemoteRiver2020a; @durandEstimatingReachaveragedDischarge2014a; @garamboisInferenceEffectiveRiver2015; @gleasonRetrievalRiverDischarge2014a; @garamboisVariationalEstimationEffective2020; @hagemannBAMBayesianAMHGManning2017a; @larnierRiverDischargeBathymetry2020a; @oubanasDischargeEstimationUngauged2018]. A specific subset of these ungauged methods are termed McFLIs or 'Mass Conserved Flow Law Inversion' algorithms [e.g. @andreadisConstrainingAssimilationSWOT2020d; @brinkerhoffConstrainingRemoteRiver2020a; @hagemannBAMBayesianAMHGManning2017a; @durandEstimatingReachaveragedDischarge2014a]. These use basic geomorphic theories and the concept of 'prior river knowledge' to estimate discharge from RS data where not all hydraulic terms are RS-able [@gleasonRemoteSensingRiver2020; @gleasonTrackingRiverFlows2017]. McFLIs are readily implemented in any river that SWOT can observe and improve our hydrological understandings of ungauged rivers when little to no information was previously available [@brinkerhoffConstrainingRemoteRiver2020a; @durandIntercomparisonRemoteSensing2016b]. Recently, McFLIs have also shown promise in providing additional and beneficial information to traditional hydrologic modeling via data assimilation of McFLIs into a traditional hydrologic model [@ishitsukaCombiningOpticalRemote2020].

While all McFLIs to date have been developed in the context of RSQ, there is no reason their logic cannot be used to estimate *k*, particularly as McFLIs often employ Bayesian inference for equifinal inverse problems. Equifinality refers to an under-constrained mathematical system that has essentially infinite parameter combinations that can produce the same result: there are in essence more unknowns than equations [@garamboisInferenceEffectiveRiver2015], and this problem is common to both RSQ and RS of *k*. Bayesian techniques have been previously used to concurrently solve for *k* and *M* in equation 2 [@applingOvercomingEquifinalityLeveraging2018; @graceFastProcessingDiel2015; @holtgrieveSimultaneousQuantificationAquatic2010]. While these studies are field based, require high-fidelity in situ data, and are not applicable to *k* estimation in ungauged rivers, they suggest that Bayesian techniques could be useful for remotely sensing *k* when parameter equifinality is a problem. This matches McFLI logic [@andreadisConstrainingAssimilationSWOT2020d].

In this context, we hypothesize that a combination of SWOT data, a Bayesian McFLI paradigm, and gas exchange theory can successfully estimate *k*. This would in turn improve our understanding of riverine gas fluxes in ungauged basins (equation 1), as well as potentially help in parameterizing stream metabolism models (equation 2). Therefore, this manuscript aims to answer two questions:

- Is ungauged RS of the gas exchange velocity (*k*) possible using soon-to-be-available SWOT river measurements?
- If possible, how will this RS method effect estimates of riverine carbon efflux from rivers?

To answer these questions, we developed a new McFLI algorithm that ingests SWOT data and produces $k_{o_2}$ estimates and its explicit Bayesian uncertainity that requires no in situ inputs of any kind (although in situ data can be ingested and will improve results). We name the RS of $k_{o_2}$ algorithm BIKER, or the 'Bayesian Inversion of the $k_{o_2}$ Evasion Rate'. We validate BIKER on simulated SWOT data for 49 SWOT-observable rivers from around the world, using hydraulic models  to produce SWOT-like data as SWOT has not yet launched. This is standard practice in the SWOT community [@frassonExploringFactorsControlling2021; @durandIntercomparisonRemoteSensing2016b]. We also quantify BIKER's sensitivity to the expected SWOT measurement errors on 17 of those rivers: while SWOT data represent a sea change in inland water monitoring, SWOT is expected to have an approximately 10cm error in water surface elevation [@biancamariaSWOTMissionIts2016], as well as river width errors [@frassonExploringFactorsControlling2021]. Finally, we use previously published dissolved $CO_2$ data to compare the bulk carbon efflux (via $CO_2$ evasion) from the 49 rivers as calculated using BIKER and previously published 'global-scope' HG models to compare BIKER with established in situ techniques.

##2 Methods
To build BIKER, we join a process-based model for $k_{o_2}$ (section 2.1) with a McFLI framework for inverting SWOT measurements (section 2.2). Following the description of that process, we describe the validation (section 2.3) and the workflow for comparing estimated bulk carbon effluxes from a suite of 'global-scope' HG models (section 2.4). A flowchart detailing the entire study is provided as Figure 1.

![Figure 1. Flowchart of methodology used in this study. We begin by choosing a process-based model for k (section 2.1) that is a function of river hydraulics measurable by SWOT. Then, we implement this model within a McFLI ('BIKER') to estimate k solely from river width and water surface slope (section 2.2). Finally, we couple BIKER with in situ CO2 data to compare BIKER-estimated gas fluxes with established in situ methods (section 2.4). See section 2.3 for the validation setup.](cache/flowchart.jpg)

###2.1 Choosing a predictive model for *k*
To predict k from just the SWOT observables,river width *W* and water surface elevation *H*, a physical model for *k* is first selected. As mentioned in section 1, there are nearly two dozen predictive models for *k* that have been devolped since the 1950s [@wangPhysicallyBasedScaling2021; @hallGasExchangeStreams2020]. Recently, many of these models were reexplored in rivers and streams by @wangPhysicallyBasedScaling2021. They significantly expanded the existing training datasets of field-measured $k_{o_2}$ by using inverse modeling of DO timeseries. This approach was also briefly mentioned in section 1. These methods calibrate $k_{o_2}$, along with *M* (discretized as seperate respiration and gross primary production terms), to high-fidelity, in situ datasets of DO with the goal of minimizing the difference between the observed DO and the modeled DO [e.g. @applingOvercomingEquifinalityLeveraging2018; @chapraDeltaMethodEstimating1991]. @wangPhysicallyBasedScaling2021 used the *streamMetabolizer* technique [@applingOvercomingEquifinalityLeveraging2018] at 35 rivers across the United States and showed that $k_{o_2}$ models deduced from the classic, process-based 'thin-film' and 'surface renewal' theories for mass transport yield approximately identical parameters when fit to field measurements or this new dataset of simulated $k_{o_2}$. They also showed through cross-validation that these models are robust to overfitting on specific sets of data (a concern with many of the aforementioned predictive models- section 1). This suggests that there are uniform scaling realtionships between certain hydraulic properties and $k_{o_2}$ that exist regardless of the training data used. See @hallGasExchangeStreams2020 and @wangPhysicallyBasedScaling2021 for more in-depth explanations of the thin-film and surface renewal theories with respect to rivers. In short, these are the standard physical models for conceptualizing gas exchange in various aquatic environments from lakes to estuaries to rivers. The better fit, and more parsimonious, of the two models from @wangPhysicallyBasedScaling2021 is reprinted as equation 3(with reported coefficients of determination $r^2$) where $\alpha$ is the fitted parameter in the linear regression. Equation 3, 2nd line was fit to 588 field measurements of $k_{o_2}$ while equation 3, third line was fit to 3,919 simulated $k_{o_2}$ values at 35 rivers.
$$\begin{eqnarray} k_{o_2} &=& \alpha U_\star \mathbf{(3)} \\ &=& 47.67U_\star (r^2:0.53) \\ &=& 49.71U_\star (r^2: 0.76)  \end{eqnarray}$$

We implement this model within BIKER (using 48 as a reasonable value for $\alpha$ given equation 3) for a few reasons. First, as suggested earlier @wangPhysicallyBasedScaling2021 provides compelling evidence that an $\alpha$ of approximately 48 is uniform across rivers and streams, regardless of the training data used. This is encouraging for use within BIKER as BIKER is specifically designed to be as river-agnostic as possible and so we sought a fully generalized equation. Second, it yields a simple linear relationship across all rivers, regardless of their size or steepeness. This is not necessairly true of other *k* models, for example those based on *eD* (section 1). Recent work has shown that *k* does not scale with *eD* via the same statistical parameters across all river geomorphologies [@ulsethDistinctAirWater2019a]. We further explored the relationship presented in that paper and found that the $k \sim eD$ relationship fundamentally breaks down in rivers with low *eD*, which corresponds to nearly all rivers that SWOT will observe (**supp figure, need to do**). Therefore, an *eD* based predictive model is less useful in SWOT rivers. Third, we sought to explicitly predict $k_{o_2}$, as its use extends beyond constraining gas fluxes from rivers and into parameterizing river metabolism models (equation 2).

We stress that most other predictive models for *k* could be implemented within BIKER. However, our goal of using BIKER on any river that is observable by SWOT necessitates that we pick an equation that is applicable across as many rivers as possible. Future work should look at using a river-specific *k* equation to pseudo-calibrate BIKER to a specific river.

###2.2 Developing BIKER
With a predictive model for $k_{o_2}$ chosen, we now implement it within the McFLI framework. The approach used here is informed by the @hagemannBAMBayesianAMHGManning2017a McFLI algorithm for ungauged RSQ, further explored in more recent work [@brinkerhoffConstrainingRemoteRiver2020a]. These papers conceptualize discharge as a Bayesian remote sensing problem, which we largely follow here to conceptualize $k_{o_2}$ as a Bayesian remote sensing problem that can be solved using SWOT data.

BIKER and Bayesian inference in general starts from Bayes rule, after acknowleding that it needs only to be specified as a proportionality to approximate the full posterior distribution (equation 4). In equation 4, $\Theta$ is some set of parameters we want to solve for (including $k_{o_2}$), *x* is the observed data, $f(x|\Theta)$ is the 'likelihood function' or sampling model conditional on the parameters, and $p(\Theta)$ is the joint prior distribution of the parameters. For BIKER, *x* is the SWOT-observables *W* and *H*. First, we detail the likelihood specification and then the prior specifications.

$$p(\Theta|x) \propto f(x|\Theta)p(\Theta) \mathbf{(4)}$$
